### 2024-11-26 pre_print/_extensions/smnnlt/sportrxiv/partials/before-body.tex
Update preprint to add PCI recommendation (and tidy front formatting)

$-- Implements the frontmatter, title page, and abstract.
$--
$-- %%%% TODO %%%%%
$-- Customize is needed, otherwise remove this partials to use Quarto default one
$-- %%%%%%%%%%%%%%%%

% set short author name
\newcounter{nauthor}
$for(by-author)$\stepcounter{nauthor}$endfor$
\ifnum\thenauthor>2
\renewcommand{\shortauthor}{$author/first$ et al. }
\else
\renewcommand{\shortauthor}{${ for(by-author) }${ by-author.name.family }${ sep } \& ${ endfor } }
\fi

\pagestyle{fancy}
\fancyhead[L]{$if(shorttitle)$$shorttitle$$else$\shortauthor$ if(year)$($year$)$else$(\the\year)$endif$$endif$}
\fancyhead[R]{}
\fancypagestyle{firstpage}{
  \fancyhf{}% clear default for head and foot
  \lhead{\href{https://sportrxiv.org}{\includegraphics[width = 40mm]{logo.png}}}
  \chead{\scriptsize{Reviewed and recommended through \href{https://healthmovsci.peercommunityin.org/articles/rec?id=44}{Peer Community in Health and Movement Sciences}}}
  \rhead{\$license-type$\vspace{0.2mm}}
  \lfoot{\color{gray}\footnotesize{All authors have read and approved this version  of the manuscript. \linebreak The manuscript was last updated on \today}}
}
\begin{flushleft}
\begin{spacing}{2.5}
\thispagestyle{firstpage}
\vspace*{0.1cm}
{\Large{\textbf{$articletitle$}}}
\end{spacing}
\vspace*{0.1cm}
$by-author:_authors.tex()[, ]$
\\
\bigskip
$by-affiliation:_affiliations.tex()[\\ ]$
\\
\bigskip
*Correspondence:
$for(by-author)$
$if(by-author.attributes.corresponding)$
\href{mailto:$by-author.email$}{\color{black}$by-author.email$}
$endif$
$endfor$
\\
\bigskip
\begin{spacing}{1}
Cite as: \shortauthor$if(year)$($year$)$else$(\the\year)$endif$. $articletitle$. \emph{SportRxiv.} DOI: 10.51224/SRXIV.348 \\
\end{spacing}
$if(suppl-link)$
  \vspace{1mm} Supplementary Materials: \href{$suppl-link$}{$suppl-link$} \\
$endif$
\vspace*{1cm}
\end{flushleft}

\newenvironment{abstractbox}
  {}

\surroundwithmdframed[
  backgroundcolor=SRXIVgreen,
  innertopmargin=10pt,
  innerbottommargin=10pt,
  linewidth=2pt,
  linecolor=gray!30,
  leftmargin=20pt,
  rightmargin=20pt,
  frametitle={\large{Abstract}},
]{abstractbox}

\begin{abstractbox}
  \footnotesize{$abstract$}
  $if(keywords)$
  Keywords: \emph{${for(keywords)}${keywords}${sep}; ${endfor}}
  $endif$
\end{abstractbox}